---
title: "WIP - Seabed Analysis"
author: "Marco Giovanni Ferrari"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
rm(list = ls(all.names = TRUE))
```

# Introduction

The aim of this code is to define an algorithm to extract the Riley
Ruggedness of the seabed under subsea cables. The ideal algorithm should
perform the following tasks:

-   Takes the **cohordinates** of a subseacable and define the
    neighbourhood around it. It should allow to work dynamically by
    defining a **radius** (in terms of kilometers) around the **cable**.

-   Once the area is defined it is necessary to compute the **Riley
    Ruggedness** in **every cell** in the **area**. Note that for every
    cell in a **spatial grid**, the Riley ruggedness is defined as
    square root of the summation of the squared difference in the
    elevation between the cell itself and its eight neighboring cells.
    In mathematical terms, a cell in row r and column c and elevation e:

    $$
    RileyRuggedness_{rc} = \sqrt{\left(  \sum_{i= r-1}^{r+1}\sum_{j= c-1}^{c+1}(e_{ij}-e_{rc})^2\right)}.
    $$

    When considering a spatial grids, cells and rows are **not indexed**
    in integer units, but have geographical coordinates, which must then
    be **ordered**.

-   Finally, following Juhász and Steinwender (2018) one can compute
    **different statistics** on the ruggedness of the cells in the area
    around the cable. In their paper, they simply considered the
    **mean** ruggedness.

In performing this analysis, the following libraries will be necessary:

```{r}
library(terra) # To import and visualize geographic data
library(dplyr) # For data manipulation
library(knitr) # For better visualization
library(ggplot2) # For data visualization and publication
library(ggthemes)
library(ggpubr)
library(purrr)
library(sf) # To manipulate geographic data and maps
```

# How bathymetry data work

First, consider the following example of bathymetry data, which can be
visualized in a $n \times 3$ dataset, with columns x (the longitude), y
(the latitude) and e (the elevation).

```{r}
geodata_example_raw <- rast("raw_data/ocean_bathymetry_data/example_bathymetry_data.tif")
geodata_example <- as.data.frame(geodata_example_raw, xy = TRUE)

geodata_example <- geodata_example %>% rename(e = example_bathymetry_data)

kable(head(geodata_example), format = "markdown")
```

Which visually appear as:

```{r, echo=TRUE,results='hide',fig.keep='all'}
ggplot() + geom_raster(data = geodata_example, aes(x=x, y=y, fill=e))+
  scale_fill_continuous_tableau() + theme_classic2()
  coord_quickmap()
```

# Extracting the area around the path

Note that, to avoid excessive calculation, we will limit this example to the area in the figure.

```{r, echo=TRUE,results='hide',fig.keep='all'}
cable_data <- st_read("raw_data/subsea_cables_location/sub_cables.geojson")

janna_raw <- cable_data %>% filter(name == "Janna")

# Limiting the analysis (and the dataset) to the area in the figure
geodata_example_border <- st_as_sfc(st_bbox(geodata_example_raw))
janna <- st_intersection(janna_raw, geodata_example_border)

ggplot() + geom_raster(data = geodata_example, aes(x=x, y=y, fill=e)) +
  geom_sf(data = janna, aes(geometry = geometry), color = "red", linewidth = 15, alpha= 0.5) +
  geom_sf(data = janna, aes(geometry = geometry), color = "red", linewidth = 1, alpha= 1) +
  scale_fill_continuous_tableau() + theme_classic2()
  coord_quickmap()

```
To compute the neighbouring area, we can leverage on the function st_buffer (package sf) and mask (package terra):

```{r, echo=TRUE,results='hide',fig.keep='all'}
# Defining the area around the cable in meters
radius = 2500
janna_area <- st_buffer(janna, dist = radius)

# Using the area around the cable to mask the dataset
janna_area <- mask(geodata_example_raw, janna_area)
janna_area <- as.data.frame(janna_area, xy = TRUE) %>% rename(e = example_bathymetry_data)

# Plotting results
ggplot() + geom_raster(data = janna_area, aes(x=x, y=y, fill=e ))+
  scale_fill_continuous_tableau() + theme_classic2()
  coord_quickmap()


```


# Computing Riley Ruggedness in each cell

First, we are going to rank each longitude and latitude, so to find the nearest cells

```{r}
janna_area <- janna_area %>% mutate(
  x_order = dense_rank(x),
  y_order = dense_rank(y)
)
```

```{r echo=TRUE}
calculate_riley_rug <- function(x_loc, y_loc, depth_loc, df) {
  # Filter neighboring points within the specified range
  neighboring_area <- df %>%
    filter(
      x_order == (x_loc + 1) | x_order == (x_loc - 1),
      y_order == (y_loc + 1) | y_order == (y_loc - 1)
    )
  
  # Check if neighboring_area has enough rows
  if (nrow(neighboring_area) < 4) {
    return(NA)  # Return NA if not enough neighbors
  }
  
  # Calculate the Riley Rug value
  riley_rug_value <- neighboring_area %>%
    mutate(
      sqhdist = (depth_loc - e)^2
    ) %>%
    summarize(riley_rug = sqrt(sum(sqhdist))) %>%
    pull(riley_rug)  # Extract riley_rug value directly
  
  return(riley_rug_value)
}
```

Applying the function:

```{r}
janna_area <- janna_area %>%
  mutate(
    riley_rug = pmap_dbl(
      list(x_order, y_order, e),
      ~ calculate_riley_rug(..1, ..2, ..3, janna_area)
    )
  )
```

Which appears as:

```{r, echo=FALSE,results='hide',fig.keep='all'}
ggplot() + geom_raster(data = janna_area, aes(x=x, y=y, fill=riley_rug)) +
  scale_fill_continuous_tableau() + theme_classic2()
  coord_quickmap()
```

